### 2017-02-08 chapters/introduction/physiology.tex

\section{Cardiac Anatomy and Physiology}

The heart is the muscular organ responsible for circulating the blood
thoughout the body.

Deoxgenated blood transported from the body trough the veins to the
right atrium (RA)

\subsection{Myocardial structure and morphology}

\subsection{Myocardial Contractility}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../main"
%%% End: