### 2017-02-08 pre_thesis/abstract.tex

\begin{abstract}
  Here is the abstract
\end{abstract}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: