### 2017-02-08 pre_thesis/acknowledgements.tex

\newenvironment{acknowledgements}%
    {\cleardoublepage\thispagestyle{empty}\null\vfill\begin{center}%
    \bfseries Acknowledgements\end{center}}%
    {\vfill\null}
        \begin{acknowledgements}
        These are the acknowledgements.
      \end{acknowledgements}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: