### 2017-03-24 chapters/introduction.tex
Update thesis to latest version

\chapter{Introduction}

% Further development in medical research require insight into complex
% systems, that is too complicated to study with mathematical models and
% super computers.

% Further development in medical sciences cannot only rely on Moore's
% law but also improvement on mathematical models and numerical
% methods.

% To lower the risk for the patient we need to more accuratley dioagnose
% the patient, and find the optimal treatment.

% Personalized medican at the level of the individual.

% Many variables to take into account

% The story of biomechanics starts in the 17th century with Giovanni
% Borelli, known as the father of Biomechanics. He was the first one to
% apply the mathematical techniques of Gallileo Galliei to study the
% movement of animals.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% THE PAST %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Over a century ago, in 1892 Robert Wood was one of the first to apply modeling
in order to understand cardiac phyiology, by approximating the ventricle as a
thinned-wall sherical shell. In this way the wall stress could be analytically
estiamated via the \emph{Law of Laplace} \cite{woods1892few}. The Law
of Laplace relates the wall stress $\sigma$ to the endocardial
pressure $P$, the endocarial radius $r$ and the wall thickness $w$,
and can in its simplest for be expressed as
\begin{align}
  \sigma = \frac{P \times r}{2 w}
\end{align}
Wall stress can be tought of as the load that the heart must work
against, and it is therefore important to understand how different
quantities affects the working heart. For instance when the heart
ejects blood, the load that the heart ejects blood against is
typically referred to as \emph{afterload}, but this is really the same
as the wall stress in that particular state. 
Even though it is obvious that the shape of a ventricle in most cases
clearly deviates from a sphere, this forumla provides clinician with a tool
that can be used to understand how diseases affect the heart.

For instance, in the case of ventricular dilation, the inner radius of
the ventricle increases, which results in increased wall stress and
hace the hear has to work harder in order to eject blood. In the other
end we have the case of concentric hypertrophy which results in wall
thickening and hence reduces the wall stress. In this case more muscle
fibers are recruited and therefore the tension experiences by each
individual unit will be less.

Since 1892 more realistic and sophisticated models has been developed
in order to better describe the anatomical as well as the biophysical
nature of the ventricle. The first extension came in 1927 to
thick-walled spheres \cite{love2013treatise}. In the 1960 the spheres
were substitued by ellipsoids, with thin-walled models in 1963
\cite{sandler1963left}, and in thick-walled models in 1968
\cite{wong1968stress}. A thourough comparison of the different early
models for ventricular wall stress is presented in
\cite{huisman1980comparison}. The first reported usage of image-based
geometries dates back to 1972 \cite{gould1972vivo} were they used
X-ray to image the left ventricle and the finite element method to
estimate the stress. The relatively poor computaional power, togther
with the low image quality really challenged the biomedical engineers
at that time, or as stated in \cite{heethaar1977computer}:
\emph{``Programming for the solution of the large number of
  simultaneous linear equations generated by the finite element method
  on computers with capacity not exceeding 32K words of memory
  required special attention to very compact storage of the
  stiffness matrix and the retrieval of its coefficients''}
A paradigm shift in the field of cardiac mechanical modeling came in
the late 1980s at the University of Auckland, and the continuum
description of the heart that were developed back
\cite{hunter1988analysis} then is still in use today.

% Looking back in time, we see that great progress have been made
% towards personalized medicin. Recent discoveries such as mapping out
% all the genes in the human genome shows that

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% THE PRESENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Today, advanced imaging techniques allow us to aquire detailed
mechanical information about the heart of the individual.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% THE FUTURE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The need for building adequate patient-specific models that captures
the geometrical information observed in the images as well as the
underlying biophysical processes is recognized as one of the key
challenges in modern bioengnieering \cite{hunter2010vision}.

In other parts of physical sciences and engineering, such as the
space industry or the weather forecast, the role of the computational
science have been cruicial for its development. Without the increasing
computational power predicted by Moore's law
\cite{brock2006understanding} and better numerical methods, such as the
finite element method, scientist would not have been able to predict
complex physical phenomena such as a Tsunami.

Despite our ability to simulate complex physical systems such as
the weather, we still have a long way to go before we are able to
simulate a human organ such as the heart. The multiscale nature of
biophysical processes has challenged bioengnieers for decades.
Even though we understand processes at each level, such as the flow of
ions in and out of the cell, or the electical transduction though the
heart, the question of how to couple these levels together is still open
\cite{noble2002modeling}.
An even greater challenge

% The problem
% of utilising this information to costrain models to become patient
% specific

% Together with increased computing power, this is perhaps
% the main reason for why computational models today are able to capture
% detailed information

\begin{itemize}
\item Building models that are specified at the level of the
  individual.
\item Using patient observations to constrain model, while at the same
  time ensuring that the underlying physical equations are satisfied. 
\end{itemize}

\input{chapters/introduction/aim_of_thesis}
\input{chapters/introduction/physiology}
\input{chapters/introduction/mechanics}
\input{chapters/introduction/patient}
\input{chapters/introduction/summary}

\pagebreak

\bibliographystyle{plain}
\bibliography{chapters/introduction/bibliography}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: