### 2017-07-20 pre_thesis/list_of_papers.tex

\section*{List of papers}

\textbf{Paper 1}
\hfill  \\ \noindent
Balaban, Gabriel, Henrik Finsberg, Hans Henrik Odland, Marie
E. Rognes, Stian Ross, Joakim Sundnes, and Samuel
Wall. \emph{High-resolution data assimilation of cardiac mechanics
  applied to a dyssynchronous ventricle.}  International journal for
numerical methods in biomedical engineering (2017). 
\\\\
\noindent
\textbf{Paper 2}
\hfill  \\ 
Finsberg, Henrik, Gabriel Balaban,  Stian Ross, Trine F. H\r{a}land,
Hans Henrik Odland, Joakim Sundnes, and Samuel
Wall. \emph{Estimating cardiac contraction through high resolution
  data assimilation of a personalized mechanical model.} Journal of
Computational Science (2017). doi:10.1016/j.jocs.2017.07.013

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: