### 2017-07-20 pre_thesis/acknowledgements.tex

% \newenvironment{acknowledgements}%
%     {\cleardoublepage\thispagestyle{empty}\null\vfill\begin{center}%
%     \bfseries Acknowledgements\end{center}}%
%     {\vfill\null}
%         \begin{acknowledgements}
%         These are the acknowledgements.
%       \end{acknowledgements}
\chapter*{Acknowledgements}
This is my thesis submitted for the degree Doctor Scientiarum,
submitted to the Department of informatics, University of Oslo. The
work conducted in this thesis took place in the period from October
2014 to September 2017.

Thanks to supervisor, funding, simula, family.

\vspace{5cm}
\noindent
Henrik Nicolay Topnes Finsberg\\
Oslo, September, 2017

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: