### 2017-10-03 pre_thesis/abstract.tex
Update figures. Change size and convert to eps

\begin{abstract}
  In this thesis we will study variational data assimilation
  techniques applied to the field of patient-specific cardiac
  mechanics modeling. We develop tenchniques for creation of
  personalized mechanical similations of human hearts, and use these
  simulations to extract biomarkers related to cardiac contractility, stress
  and work. The tools developed in this thesis could potentially be used to assess
  additional mechanical information \emph{in vivo}, beyond the
  capabilities of modern imaging techniques.   
\end{abstract}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: