### 2017-10-03 pre_thesis/acknowledgements.tex

\section*{Acknowledgements}

This is my thesis submitted for the degree Doctor Scientiarum,
submitted to the Department of Informatics, University of Oslo. The
work conducted in this thesis took place in the period from October
2014 to September 2017.\\
\indent First of all, I wish to deeply acknowledge the Center for Cardiological
Innovation (CCI) for providing the financial support for my PhD.\\
\indent I would especially like thank my main supervisors Joakim and Sam,
which both have been excellent in providing useful feedback, and
directions for my thesis. Joakim, thank you for allways taking time to
answer any question I may have, whether it is related to work or
private. Sam, thank you for setting up all the collaborations and for
guiding my work in the right direction.\\
\indent I would also like to thank Simula Research Laboratory for providing
an outstanding working environment. In my opinion, the including
management and the nice people at Simula, makes Simula the perfect
place to pursue a PhD. A special thank to the people from the Cardiac
Modeling group, in particular to Siri which I shared office with for most
of my PhD, and to Gabriel who I worked very close with during my first
year. \\
\indent I am also grateful to the CCI for giving me the opportunity to work
with, and learn from such exceptional medical researchers and for
providing an environment for interdisciplinary collaborations. In
particular I would like to thank, Hans Henrik, John and Espen for many
interesting and fruitful discussions. \\
\indent Finally, I would like to thank my family for always reminding me what
truly matters in life. Elisabeth, I am
deeply grateful for all the support you have given me during these years. 
% \end{abstract}

\vspace{5cm}
\noindent
Henrik Nicolay Topnes Finsberg\\
Oslo, September, 2017

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: