### 2017-10-03 pre_thesis/list_of_papers.tex

\section*{List of papers}

\textbf{Paper 1}
\hfill  \\ \noindent
Gabriel Balban, Henrik Finsberg, Hans Henrik Odland, Marie
E. Rognes, Stian Ross, Joakim Sundnes, and Samuel
Wall. \emph{High-resolution data assimilation of cardiac mechanics
  applied to a dyssynchronous ventricle.}  International journal for
numerical methods in biomedical engineering
(2017). doi:10.1002/cnm.2863
Status: published
\\\\
\noindent
\textbf{Paper 2} 
\hfill  \\ 
Henrik Finsberg, Gabriel Balaban,  Stian Ross, Trine F. H\r{a}land,
Hans Henrik Odland, Joakim Sundnes, and Samuel
Wall. \emph{Estimating cardiac contraction through high resolution
  data assimilation of a personalized mechanical model.} Journal of
Computational Science (2017). doi:10.1016/j.jocs.2017.07.013
\\Status: published
\\\\
\noindent
\textbf{Paper 3}
\hfill  \\
Henrik Finsberg, Ce Xi,  Ju Le Tan,  Liang Zhong, Martin Genet, Joakim
Sundnes, Lik Chuan Lee, and Samuel T. Wall.
\emph{Personalized bi-ventricular mechanical analysis using
  gradient-based optimization.} International journal for
numerical methods in biomedical engineering
\\Status: submitted
\\\\
\noindent
\textbf{Paper 4}
\hfill  \\
Henrik Finsberg, John Aalen,Camilla K.
Larsen, Joakim Sundnes, Otto A. Smiseth, and 
Samuel T. Wall.
\emph{Assessment of regional myocardial work from a
patient-specific cardiac mechanics model.} 
\\Status: to be submitted

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: